---
title: "Examples of working with the FARS data"
author: "Matthew R. MacLeod"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples of working with the FARS data using Week4MRM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


The Week4MRM pacakge provides functions for working with files downloaded from the Fatality Analysis Reporting System (FARS) at https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars, allowing data to be summarized and plotted on a map.

This package relies on the existence of files with names of the form 

```{r eval=FALSE}
accident_NNNN.csv
```

or

```{r eval=FALSE}
accident_NNNN.csv.bz2
```

Where **NNNN** is a year.  


Other than suppporting utility functions, the most basic funciton of use to the user is to read in one or more years worth of data to a, using the following syntax:

```{r eval=FALSE}
fars_read_years(2014)
fars_read_years(c(2014,2015,2016))
```


More useful is a function which creates a table of the month and year of each incident in the FARS database
for one or more specified years, then computes and returns a summary table of the number of incidents
with the months in rows and years in columns:

```{r eval=FALSE}
fars_summarize_years(2014)
fars_summarize_years(c(2014,2015,2016)
```

Finally, there is a function which plots the locations of FARS incidents in a given state in a given year. States are nubmered in alphabetical order (see p. 50 of the guide at the following link: https://crashstats.nhtsa.dot.gov/Api/Public/ViewPublication/812449), including District of Columbia, Peurto Rico and Virgin Islands, as entries, but skipping the numbers 3, 7 and 14.  Example invocations are as follows:

```{r eval=FALSE}
fars_map_state(36,2013)
fars_map_state(34,2014)
fars_map_state(34,2014)
```
